---
title: "Walking SALURBAL-Climate Proposal"
author:
  - name: Amanda Silva Magalhães
    affiliation: Federal University of Minas Gerais - Brazil
  - name: Waleska Teixeira Caiaffa
    affiliation: Federal University of Minas Gerais - Brazil

date: today

bibliography: references.bib
---


# Setup

```{r}
#| label: Clean environment
#| code-summary: Clean environment

rm(list = ls())
```

```{r}
#| label: Directory
#| code-summary: Directory

setwd("C:/Users/amand/Amanda/GitHub/walking")
```

```{r}
#| label: Libraries
#| code-summary: Libraries

pacman::p_load(# Data import
               haven, readxl,
               # Data manipulation
               dplyr, tidyr, stringr, purrr,
               # Data visualization
               ggplot2, patchwork, ggeffects, grid, magick,
               # Geospatial
               sf, ggspatial, tidygeocoder, rnaturalearth,
               # Statistical modeling
               splines, glmmTMB, lme4, mgcv,
               # Post-model-fitting
               performance, DHARMa,
               # Correlation
               Hmisc, reshape2,
               # Parallel computing
               parallel, pbapply
)
```

```{r}
#| label: Data
#| code-summary: Data

dta <- haven::read_dta("walking.dta")
```


# Distribution of outcomes

Global walking levels differed across countries. Median walking was higher in Mexico and El Salvador and lower in Colombia and Guatemala. Greater variability was observed in Mexico and El Salvador, while extreme values were present in all countries.

```{r}
#| label: Global
#| code-summary: Global

dta %>%
  mutate(
    iso2 = recode(
      iso2,
      CO = "Colombia",
      GT = "Guatemala",
      SV = "El Salvador",
      MX = "Mexico"
    )
  ) %>%
  ggplot(aes(x = iso2, y = svypawalk, fill = iso2)) +
  geom_boxplot() +
  scale_fill_manual(
    values = c(
      "Colombia" = "#F0BF21",
      "Guatemala" = "#BCD5FF",
      "El Salvador" = "#303B56",
      "Mexico" = "#238179"
    )
  ) +
  scale_y_continuous(
    breaks = seq(0, 1500, by = 250)
  ) +
  labs(
    x = "",
    y = "Global walking (min/week) \n"
  ) +
  theme_minimal(base_size = 18) +
  theme(
    legend.position = "none"
  )
```

Leisure walking levels were low in both countries, with median values close to zero. Greater variability and higher maximum values were observed in Colombia than in Guatemala.

```{r}
#| label: Leisure
#| code-summary: Leisure

dta %>%
  filter(iso2 %in% c("CO", "GT")) %>%
  mutate(
    iso2 = recode(
      iso2,
      CO = "Colombia",
      GT = "Guatemala"
    )
  ) %>%
  ggplot(aes(x = iso2, y = svypawlei, fill = iso2)) +
  geom_boxplot() +
  scale_fill_manual(
    values = c(
      "Colombia" = "#F0BF21",
      "Guatemala" = "#BCD5FF"
    )
  ) +
  scale_y_continuous(
    breaks = seq(0, 1500, by = 250)
  ) +
  labs(
    x = "",
    y = "Leisure walking (min/week) \n"
  ) +
  theme_minimal(base_size = 18) +
  theme(
    legend.position = "none"
  )
```

Transportation walking levels were higher in Colombia than in Guatemala, as indicated by a higher median value. Greater variability and higher maximum values were also observed in Colombia.

```{r}
#| label: Transportation
#| code-summary: Transportation

dta %>%
  filter(iso2 %in% c("CO", "GT")) %>%
  mutate(
    iso2 = recode(
      iso2,
      CO = "Colombia",
      GT = "Guatemala"
    )
  ) %>%
  ggplot(aes(x = iso2, y = svypawtrn, fill = iso2)) +
  geom_boxplot() +
  scale_fill_manual(
    values = c(
      "Colombia" = "#F0BF21",
      "Guatemala" = "#BCD5FF"
    )
  ) +
  scale_y_continuous(
    breaks = seq(0, 1500, by = 250)
  ) +
  labs(
    x = "",
    y = "Transportation walking (min/week) \n"
  ) +
  theme_minimal(base_size = 18) +
  theme(
    legend.position = "none"
  )
```

# Distribution of exposure

The distribution of mean temperature during the week preceding the interview varied across countries. Guatemala and El Salvador showed relatively narrow temperature ranges, whereas Colombia and Mexico exhibited greater variability and a wider distribution of temperatures. Histogram frequencies represent the number of individuals interviewed within each temperature interval.

```{r}
#| label: Histogram
#| code-summary: Histogram

dta %>%
  mutate(
    iso2 = recode(
      iso2,
      CO = "Colombia",
      GT = "Guatemala",
      SV = "El Salvador",
      MX = "Mexico"
    )
  ) %>%
  ggplot(aes(x = l1adtemp_pw_1_7_mean)) +
  geom_histogram(
    binwidth = 1,
    fill = "#4666A0",
    color = "white"
  ) +
  facet_wrap(~ iso2, scales = "free_y") +
  scale_x_continuous(
    breaks = seq(0, 35, by = 5)
  ) +
  labs(
    x = "\n Mean temperature (°C)",
    y = "Frequency \n"
  ) +
  theme_minimal(base_size = 18)
```

```{r}
#| label: Boxplot
#| code-summary: Boxplot

dta %>%
  mutate(
    iso2 = recode(
      iso2,
      CO = "Colombia",
      GT = "Guatemala",
      SV = "El Salvador",
      MX = "Mexico"
    )
  ) %>%
  ggplot(aes(x = iso2, y = l1adtemp_pw_1_7_mean, fill = iso2)) +
  geom_boxplot() +
  scale_fill_manual(
    values = c(
      "Colombia" = "#F0BF21",
      "Guatemala" = "#BCD5FF",
      "El Salvador" = "#303B56",
      "Mexico" = "#238179"
    )
  ) +
  scale_y_continuous(
    breaks = seq(0, 35, by = 5)
  ) +
  labs(
    x = "",
    y = "Mean temperature (°C) \n"
  ) +
  theme_minimal(base_size = 18) +
  theme(
    legend.position = "none"
  )
```

# Outcome by exposure

All walking domains showed a non-linear association with mean temperature during the week preceding the interview. These exploratory analyses were restricted to participants reporting more than 0 min/week of walking to focus on variation among individuals who engaged in walking activities.

```{r}
#| label: Global by temperature
#| code-summary: Global by temperature

dta %>%
  filter(svypawalk > 0) %>%
  ggplot(
    aes(
      x = l1adtemp_pw_1_7_mean,
      y = svypawalk
    )
  ) +
  geom_smooth(se = TRUE) +
  labs(
    x = "\n Mean temperature (°C)",
    y = "Global walking (min/week) \n"
  ) +
  theme_minimal(base_size = 18)
```

```{r}
#| label: Leisure by temperature
#| code-summary: Leisure by temperature

dta %>%
  filter(svypawlei > 0) %>%
  ggplot(
    aes(
      x = l1adtemp_pw_1_7_mean,
      y = svypawlei
    )
  ) +
  geom_smooth(se = TRUE) +
  labs(
    x = "\n Mean temperature (°C)",
    y = "Leisure walking (min/week) \n"
  ) +
  theme_minimal(base_size = 18)
```

```{r}
#| label: Transportation by temperature
#| code-summary: Transportation by temperature

dta %>%
  filter(svypawtrn > 0) %>%
  ggplot(
    aes(
      x = l1adtemp_pw_1_7_mean,
      y = svypawtrn
    )
  ) +
  geom_smooth(se = TRUE) +
  labs(
    x = "\n Mean temperature (°C)",
    y = "Transportation walking (min/week) \n"
  ) +
  theme_minimal(base_size = 18)
```

