---
title: "Walking SALURBAL-Climate Proposal"
author:
  - name: Amanda Silva Magalhães
    affiliation: Federal University of Minas Gerais - Brazil
  - name: Waleska Teixeira Caiaffa
    affiliation: Federal University of Minas Gerais - Brazil

date: today

bibliography: references.bib
---


# Setup

```{r}
#| label: Clean environment
#| code-summary: Clean environment

rm(list = ls())
```

```{r}
#| label: Directory
#| code-summary: Directory

setwd("C:/Users/amand/Amanda/GitHub/walking")
```

```{r}
#| label: Libraries
#| code-summary: Libraries

pacman::p_load(
  haven, readxl,
  dplyr, tidyr, stringr,
  ggplot2, patchwork,
  gt,
  sf,
  splines, glmmTMB, lme4, mgcv,
  performance, DHARMa,
  corrplot
)
```

```{r}
#| label: Data
#| code-summary: Data

dta <- haven::read_dta("walking.dta")
```


# Distribution of outcomes

Global walking levels differed across countries. Median walking was higher in Mexico and El Salvador and lower in Colombia and Guatemala. Greater variability was observed in Mexico and El Salvador, while extreme values were present in all countries.

```{r}
#| label: Global
#| code-summary: Global

dta %>%
  mutate(
    iso2 = recode(
      iso2,
      CO = "Colombia",
      GT = "Guatemala",
      SV = "El Salvador",
      MX = "Mexico"
    )
  ) %>%
  ggplot(aes(x = iso2, y = svypawalk, fill = iso2)) +
  geom_boxplot() +
  scale_fill_manual(
    values = c(
      "Colombia" = "#F0BF21",
      "Guatemala" = "#BCD5FF",
      "El Salvador" = "#303B56",
      "Mexico" = "#238179"
    )
  ) +
  scale_y_continuous(
    breaks = seq(0, 1500, by = 250)
  ) +
  labs(
    x = "",
    y = "Global walking (min/week) \n"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "none"
  )
```

Leisure walking levels were low in both countries, with median values close to zero. Greater variability and higher maximum values were observed in Colombia than in Guatemala.

```{r}
#| label: Leisure
#| code-summary: Leisure

dta %>%
  filter(iso2 %in% c("CO", "GT")) %>%
  mutate(
    iso2 = recode(
      iso2,
      CO = "Colombia",
      GT = "Guatemala"
    )
  ) %>%
  ggplot(aes(x = iso2, y = svypawlei, fill = iso2)) +
  geom_boxplot() +
  scale_fill_manual(
    values = c(
      "Colombia" = "#F0BF21",
      "Guatemala" = "#BCD5FF"
    )
  ) +
  scale_y_continuous(
    breaks = seq(0, 1500, by = 250)
  ) +
  labs(
    x = "",
    y = "Leisure walking (min/week) \n"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "none"
  )
```

Transportation walking levels were higher in Colombia than in Guatemala, as indicated by a higher median value. Greater variability and higher maximum values were also observed in Colombia.

```{r}
#| label: Transportation
#| code-summary: Transportation

dta %>%
  filter(iso2 %in% c("CO", "GT")) %>%
  mutate(
    iso2 = recode(
      iso2,
      CO = "Colombia",
      GT = "Guatemala"
    )
  ) %>%
  ggplot(aes(x = iso2, y = svypawtrn, fill = iso2)) +
  geom_boxplot() +
  scale_fill_manual(
    values = c(
      "Colombia" = "#F0BF21",
      "Guatemala" = "#BCD5FF"
    )
  ) +
  scale_y_continuous(
    breaks = seq(0, 1500, by = 250)
  ) +
  labs(
    x = "",
    y = "Transportation walking (min/week) \n"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "none"
  )
```

# Distribution of exposure

The distribution of mean temperature during the week preceding the interview varied across countries. Guatemala and El Salvador showed relatively narrow temperature ranges, whereas Colombia and Mexico exhibited greater variability and a wider distribution of temperatures. Histogram frequencies represent the number of individuals interviewed within each temperature interval.

```{r}
#| label: Histogram
#| code-summary: Histogram

dta %>%
  mutate(
    iso2 = recode(
      iso2,
      CO = "Colombia",
      GT = "Guatemala",
      SV = "El Salvador",
      MX = "Mexico"
    )
  ) %>%
  ggplot(aes(x = l1adtemp_pw_1_7_mean)) +
  geom_histogram(
    binwidth = 1,
    fill = "#4666A0",
    color = "white"
  ) +
  facet_wrap(~ iso2, scales = "free_y") +
  scale_x_continuous(
    breaks = seq(0, 35, by = 5)
  ) +
  labs(
    x = "\n Mean temperature (°C)",
    y = "Frequency \n"
  ) +
  theme_minimal(base_size = 14)
```

```{r}
#| label: Boxplot
#| code-summary: Boxplot

dta %>%
  mutate(
    iso2 = recode(
      iso2,
      CO = "Colombia",
      GT = "Guatemala",
      SV = "El Salvador",
      MX = "Mexico"
    )
  ) %>%
  ggplot(aes(x = iso2, y = l1adtemp_pw_1_7_mean, fill = iso2)) +
  geom_boxplot() +
  scale_fill_manual(
    values = c(
      "Colombia" = "#F0BF21",
      "Guatemala" = "#BCD5FF",
      "El Salvador" = "#303B56",
      "Mexico" = "#238179"
    )
  ) +
  scale_y_continuous(
    breaks = seq(0, 35, by = 5)
  ) +
  labs(
    x = "",
    y = "Mean temperature (°C) \n"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "none"
  )
```

# Outcome by exposure

All walking domains showed a non-linear association with mean temperature during the week preceding the interview. These exploratory analyses were restricted to participants reporting more than 0 min/week of walking to focus on variation among individuals who engaged in walking activities.

```{r}
#| label: Global by temperature
#| code-summary: Global by temperature

dta %>%
  filter(svypawalk > 0) %>%
  ggplot(
    aes(
      x = l1adtemp_pw_1_7_mean,
      y = svypawalk
    )
  ) +
  geom_smooth(se = TRUE) +
  labs(
    x = "\n Mean temperature (°C)",
    y = "Global walking (min/week) \n"
  ) +
  theme_minimal(base_size = 14)
```

```{r}
#| label: Leisure by temperature
#| code-summary: Leisure by temperature

dta %>%
  filter(svypawlei > 0) %>%
  ggplot(
    aes(
      x = l1adtemp_pw_1_7_mean,
      y = svypawlei
    )
  ) +
  geom_smooth(se = TRUE) +
  labs(
    x = "\n Mean temperature (°C)",
    y = "Leisure walking (min/week) \n"
  ) +
  theme_minimal(base_size = 14)
```

```{r}
#| label: Transportation by temperature
#| code-summary: Transportation by temperature

dta %>%
  filter(svypawtrn > 0) %>%
  ggplot(
    aes(
      x = l1adtemp_pw_1_7_mean,
      y = svypawtrn
    )
  ) +
  geom_smooth(se = TRUE) +
  labs(
    x = "\n Mean temperature (°C)",
    y = "Transportation walking (min/week) \n"
  ) +
  theme_minimal(base_size = 14)
```

# Tabel 1

```{r}
#| label: Functions
#| code-summary: Functions

make_continuous_table <- function(data, variable, label, digits = 1) {

  safe_min <- function(x) {
    if (all(is.na(x))) NA_real_ else min(x, na.rm = TRUE)
  }

  safe_max <- function(x) {
    if (all(is.na(x))) NA_real_ else max(x, na.rm = TRUE)
  }

  total <- data.frame(
    iso2 = "Total",
    median = median(data[[variable]], na.rm = TRUE),
    Q1_val = quantile(data[[variable]], .25, na.rm = TRUE),
    Q3_val = quantile(data[[variable]], .75, na.rm = TRUE),
    min_val = safe_min(data[[variable]]),
    max_val = safe_max(data[[variable]])
  )

  by_iso2 <- data %>%
    filter(!is.na(iso2)) %>%
    group_by(iso2) %>%
    summarise(
      median = median(.data[[variable]], na.rm = TRUE),
      Q1_val = quantile(.data[[variable]], .25, na.rm = TRUE),
      Q3_val = quantile(.data[[variable]], .75, na.rm = TRUE),
      min_val = safe_min(.data[[variable]]),
      max_val = safe_max(.data[[variable]]),
      .groups = "drop"
    )

  stats <- bind_rows(total, by_iso2)

  bind_rows(

    stats %>%
      mutate(
        value = ifelse(
          is.na(median),
          NA_character_,
          paste0(
            format(round(median, digits), trim = TRUE),
            " (",
            format(round(Q1_val, digits), trim = TRUE),
            "; ",
            format(round(Q3_val, digits), trim = TRUE),
            ")"
          )
        ),
        Characteristic = paste0(label, ", median (Q1; Q3)")
      ) %>%
      select(Characteristic, iso2, value),

    stats %>%
      mutate(
        value = ifelse(
          is.na(min_val),
          NA_character_,
          paste0(
            format(round(min_val, digits), trim = TRUE),
            " - ",
            format(round(max_val, digits), trim = TRUE)
          )
        ),
        Characteristic = paste0(label, ", min-max")
      ) %>%
      select(Characteristic, iso2, value)

  )
}

make_categorical_table <- function(data, variable, label) {

  total <- data %>%
    filter(!is.na(.data[[variable]])) %>%
    count(.data[[variable]]) %>%
    mutate(
      iso2 = "Total",
      percent = 100 * n / sum(n),
      value = paste0(
        format(n, big.mark = ","),
        " (",
        formatC(percent, format = "f", digits = 1),
        ")"
      )
    ) %>%
    rename(level = all_of(variable)) %>%
    select(iso2, level, value)

  by_iso2 <- data %>%
    filter(!is.na(.data[[variable]]), !is.na(iso2)) %>%
    group_by(iso2, .data[[variable]]) %>%
    summarise(
      n = n(),
      .groups = "drop"
    ) %>%
    group_by(iso2) %>%
    mutate(
      percent = 100 * n / sum(n),
      value = paste0(
        format(n, big.mark = ","),
        " (",
        formatC(percent, format = "f", digits = 1),
        ")"
      )
    ) %>%
    ungroup() %>%
    rename(level = all_of(variable)) %>%
    select(iso2, level, value)

  bind_rows(total, by_iso2) %>%
    mutate(
      level = as.character(level),
      Characteristic = paste0("   ", label, " - ", level)
    ) %>%
    select(Characteristic, iso2, value) %>%
    bind_rows(
      tibble(
        Characteristic = label,
        iso2 = "Total",
        value = ""
      ),
      .
    )
}
```

```{r}
#| label: Context
#| code-summary: Context

# Number of cities

n_cities <- dta %>%
  filter(salid1_tag == 1) %>%
  count(iso2, name = "value") %>%
  bind_rows(
    summarise(., iso2 = "Total", value = sum(value))
  ) %>%
  mutate(
    value = format(
      value,
      big.mark = ",",
      decimal.mark = ".",
      scientific = FALSE
    ),
    Characteristic = "Number of cities"
  ) %>%
  select(Characteristic, iso2, value)

# Participants per city

participants_city <- make_continuous_table(
  data = dta %>% filter(!is.na(salid1)),
  variable = "salid1_obs",
  label = "Participants per city"
)

# Number of sub-cities

n_subcities <- dta %>%
  filter(salid2_tag == 1) %>%
  count(iso2, name = "value") %>%
  bind_rows(
    summarise(., iso2 = "Total", value = sum(value))
  ) %>%
  mutate(
    value = format(
      value,
      big.mark = ",",
      decimal.mark = "."
    ),
    Characteristic = "Number of sub-cities"
  ) %>%
  select(Characteristic, iso2, value)

# Participants per sub-city

participants_subcity <- make_continuous_table(
  data = dta %>% filter(!is.na(salid2)),
  variable = "salid2_obs",
  label = "Participants per sub-city"
)

# Number of participants

n_participants <- dta %>%
  count(iso2, name = "n") %>%
  bind_rows(
    summarise(., iso2 = "Total", n = sum(n))
  ) %>%
  mutate(
    pct = round(100 * n / n[iso2 == "Total"], 1),
    value = ifelse(
      iso2 == "Total",
      paste0(format(n, big.mark = ","), " (100.0)"),
      paste0(format(n, big.mark = ","), " (", pct, ")")
    ),
    Characteristic = "Number of participants, n (%)"
  ) %>%
  select(Characteristic, iso2, value)

# Context table

context_table <- bind_rows(
  n_cities,
  participants_city,
  n_subcities,
  participants_subcity,
  n_participants
)

context_table <- context_table %>%
  mutate(Domain = "Context")
```

```{r}
#| label: Individual-level
#| code-summary: Individual-level

# Recoding categorical variables

dta <- dta %>%
  mutate(

    svymale = factor(
      svymale,
      levels = c(0, 1),
      labels = c("Female", "Male")
    ),

    svyedu = case_when(
      svyedu %in% c(9998, 9999) ~ NA_real_,
      TRUE ~ as.numeric(svyedu)
    ),

    svyedu = factor(
      svyedu,
      levels = c(1, 2, 3, 4),
      labels = c(
        "Less than primary",
        "Primary",
        "Secondary",
        "University or higher"
      )
    ),

    svypawalkcat = factor(
      svypawalkcat,
      levels = c(0, 1),
      labels = c("Inactive", "Active")
    ),

    svypawleicat = factor(
      svypawleicat,
      levels = c(0, 1),
      labels = c("Inactive", "Active")
    ),

    svypawtrncat = factor(
      svypawtrncat,
      levels = c(0, 1),
      labels = c("Inactive", "Active")
    )
  )

# Age

age_total <- dta %>%
  summarise(
    iso2 = "Total",
    mean = mean(svyage, na.rm = TRUE),
    sd = sd(svyage, na.rm = TRUE),
    min_val = min(svyage, na.rm = TRUE),
    max_val = max(svyage, na.rm = TRUE)
  )

age_by_iso2 <- dta %>%
  group_by(iso2) %>%
  summarise(
    mean = mean(svyage, na.rm = TRUE),
    sd = sd(svyage, na.rm = TRUE),
    min_val = min(svyage, na.rm = TRUE),
    max_val = max(svyage, na.rm = TRUE),
    .groups = "drop"
  )

age_stats <- bind_rows(age_total, age_by_iso2)

age_table <- bind_rows(

  age_stats %>%
    mutate(
      value = paste0(
        format(round(mean, 1), trim = TRUE),
        " (",
        format(round(sd, 1), trim = TRUE),
        ")"
      ),
      Characteristic = "Age, mean (SD)"
    ) %>%
    select(Characteristic, iso2, value),

  age_stats %>%
    mutate(
      value = paste0(
        format(round(min_val, 1), trim = TRUE),
        " - ",
        format(round(max_val, 1), trim = TRUE)
      ),
      Characteristic = "Age, min-max"
    ) %>%
    select(Characteristic, iso2, value)

)

# Sex

sex_table <- make_categorical_table(
  dta,
  "svymale",
  "Sex, n (%)"
)

# Educational level

education_table <- make_categorical_table(
  dta,
  "svyedu",
  "Educational level, n (%)"
)

# Global walking

global_walk <- make_continuous_table(
  dta,
  "svypawalk",
  "Global walking (min/week)",
  digits = 0
)

global_walk_cat <- make_categorical_table(
  dta,
  "svypawalkcat",
  "Global walking, n (%)"
)

# Leisure walking

leisure_walk <- make_continuous_table(
  dta,
  "svypawlei",
  "Leisure walking (min/week)",
  digits = 0
)

leisure_walk_cat <- make_categorical_table(
  dta,
  "svypawleicat",
  "Leisure walking, n (%)"
)

# Transportation walking

transport_walk <- make_continuous_table(
  dta,
  "svypawtrn",
  "Transportation walking (min/week)",
  digits = 0
)

transport_walk_cat <- make_categorical_table(
  dta,
  "svypawtrncat",
  "Transportation walking, n (%)"
)

# Final individual-level table

individual_table <- bind_rows(
  age_table,
  sex_table,
  education_table,
  global_walk,
  global_walk_cat,
  leisure_walk,
  leisure_walk_cat,
  transport_walk,
  transport_walk_cat
) %>%
  mutate(
    Domain = "Individual-level"
  )
```

```{r}
#| label: Natural environment
#| code-summary: Natural environment

# Mean temperature

temperature_city <- dta %>%
  filter(!is.na(salid1)) %>%
  group_by(iso2, salid1) %>%
  summarise(
    value = mean(l1adtemp_pw_1_7_mean, na.rm = TRUE),
    .groups = "drop"
  )

temperature_total <- temperature_city %>%
  summarise(
    iso2 = "Total",
    median = median(value, na.rm = TRUE),
    P10 = quantile(value, .10, na.rm = TRUE),
    P90 = quantile(value, .90, na.rm = TRUE),
    min_val = min(value, na.rm = TRUE),
    max_val = max(value, na.rm = TRUE)
  )

temperature_by_iso2 <- temperature_city %>%
  group_by(iso2) %>%
  summarise(
    median = median(value, na.rm = TRUE),
    P10 = quantile(value, .10, na.rm = TRUE),
    P90 = quantile(value, .90, na.rm = TRUE),
    min_val = min(value, na.rm = TRUE),
    max_val = max(value, na.rm = TRUE),
    .groups = "drop"
  )

temperature_stats <- bind_rows(
  temperature_total,
  temperature_by_iso2
)

temperature_table <- bind_rows(

  temperature_stats %>%
    mutate(
      value = paste0(
        format(round(median, 1), trim = TRUE),
        " (",
        format(round(P10, 1), trim = TRUE),
        "; ",
        format(round(P90, 1), trim = TRUE),
        ")"
      ),
      Characteristic = "Mean temperature (°C), median (P10; P90)"
    ) %>%
    select(Characteristic, iso2, value),

  temperature_stats %>%
    mutate(
      value = paste0(
        format(round(min_val, 1), trim = TRUE),
        " - ",
        format(round(max_val, 1), trim = TRUE)
      ),
      Characteristic = "Mean temperature (°C), min-max"
    ) %>%
    select(Characteristic, iso2, value)

)

# Climate zone

dta <- dta %>%
  mutate(
    becczl1adcat = as.character(
      haven::as_factor(becczl1adcat)
    )
  )

climate_city <- dta %>%
  filter(
    !is.na(becczl1adcat),
    !is.na(salid1)
  ) %>%
  distinct(
    iso2,
    salid1,
    becczl1adcat
  )

climate_total <- climate_city %>%
  count(becczl1adcat) %>%
  mutate(
    iso2 = "Total",
    percent = 100 * n / sum(n),
    value = paste0(
      format(n, big.mark = ","),
      " (",
      formatC(percent, digits = 1, format = "f"),
      ")"
    )
  ) %>%
  transmute(
    Characteristic = paste0("   ", becczl1adcat),
    iso2,
    value
  )

climate_by_iso2 <- climate_city %>%
  count(
    iso2,
    becczl1adcat
  ) %>%
  group_by(iso2) %>%
  mutate(
    percent = 100 * n / sum(n),
    value = paste0(
      format(n, big.mark = ","),
      " (",
      formatC(percent, digits = 1, format = "f"),
      ")"
    )
  ) %>%
  ungroup() %>%
  transmute(
    Characteristic = paste0("   ", becczl1adcat),
    iso2,
    value
  )

climate_table <- bind_rows(
  tibble(
    Characteristic = "Major climate zone, n (%)",
    iso2 = "Total",
    value = ""
  ),
  climate_total,
  climate_by_iso2
)

# Greenness

greenness_subcity <- dta %>%
  filter(!is.na(salid2)) %>%
  group_by(
    iso2,
    salid2
  ) %>%
  summarise(
    value = mean(becmedndvinwl2, na.rm = TRUE),
    .groups = "drop"
  )

greenness_table <- make_continuous_table(
  greenness_subcity,
  "value",
  "Greenness",
  digits = 1
)

# Final natural environment table

natural_table <- bind_rows(
  temperature_table,
  climate_table,
  greenness_table
)

natural_table <- natural_table %>%
  mutate(Domain = "Natural environment")
```

```{r}
#| label: Built environment
#| code-summary: Built environment

# Presence of a BRT

brt_city <- dta %>%
  filter(
    !is.na(becprsbrtl1ad),
    !is.na(salid1)
  ) %>%
  distinct(
    iso2,
    salid1,
    becprsbrtl1ad
  )

brt_total <- brt_city %>%
  count(becprsbrtl1ad) %>%
  mutate(
    iso2 = "Total",
    percent = 100 * n / sum(n),
    value = paste0(
      format(n, big.mark = ","),
      " (",
      formatC(percent, digits = 1, format = "f"),
      ")"
    )
  ) %>%
  transmute(
    Characteristic = paste0("   ", becprsbrtl1ad),
    iso2,
    value
  )

brt_by_iso2 <- brt_city %>%
  count(
    iso2,
    becprsbrtl1ad
  ) %>%
  group_by(iso2) %>%
  mutate(
    percent = 100 * n / sum(n),
    value = paste0(
      format(n, big.mark = ","),
      " (",
      formatC(percent, digits = 1, format = "f"),
      ")"
    )
  ) %>%
  ungroup() %>%
  transmute(
    Characteristic = paste0("   ", becprsbrtl1ad),
    iso2,
    value
  )

brt_table <- bind_rows(
  tibble(
    Characteristic = "Presence of a BRT, n (%)",
    iso2 = "Total",
    value = ""
  ),
  brt_total,
  brt_by_iso2
)

# Percentage of urban area

urban_area_city <- dta %>%
  filter(!is.na(salid1)) %>%
  group_by(
    iso2,
    salid1
  ) %>%
  summarise(
    value = mean(becpcturbanl1ad, na.rm = TRUE),
    .groups = "drop"
  )

urban_area_table <- make_continuous_table(
  urban_area_city,
  "value",
  "Percentage of urban area",
  digits = 1
)

# Population density

population_density_subcity <- dta %>%
  filter(!is.na(salid2)) %>%
  group_by(
    iso2,
    salid2
  ) %>%
  summarise(
    value = mean(becpopdensl2, na.rm = TRUE),
    .groups = "drop"
  )

population_density_table <- make_continuous_table(
  population_density_subcity,
  "value",
  "Population density",
  digits = 0
)

# Intersection density

intersection_density_subcity <- dta %>%
  filter(!is.na(salid2)) %>%
  group_by(
    iso2,
    salid2
  ) %>%
  summarise(
    value = mean(becadintdensl2, na.rm = TRUE),
    .groups = "drop"
  )

intersection_density_table <- make_continuous_table(
  intersection_density_subcity,
  "value",
  "Intersection density",
  digits = 1
)

# Parks density

parks_density_subcity <- dta %>%
  filter(!is.na(salid2)) %>%
  group_by(
    iso2,
    salid2
  ) %>%
  summarise(
    value = mean(becparknumpercapl2, na.rm = TRUE),
    .groups = "drop"
  )

parks_density_table <- make_continuous_table(
  parks_density_subcity,
  "value",
  "Parks density",
  digits = 1
)

# Final built environment table

built_table <- bind_rows(
  brt_table,
  urban_area_table,
  population_density_table,
  intersection_density_table,
  parks_density_table
)

built_table <- built_table %>%
  mutate(Domain = "Built environment")
```

```{r}
#| label: Social environment
#| code-summary: Social environment

# Social environment index

social_city <- dta %>%
  filter(!is.na(salid1)) %>%
  group_by(
    iso2,
    salid1
  ) %>%
  summarise(
    value = mean(cnssei_l1ad, na.rm = TRUE),
    .groups = "drop"
  )

social_table <- make_continuous_table(
  social_city,
  "value",
  "Social environment index",
  digits = 1
)

social_table <- social_table %>%
  mutate(Domain = "Social environment")
```

```{r}
#| label: Table 1
#| code-summary: Table 1

library(gt)

table1 <- bind_rows(
  context_table,
  individual_table,
  natural_table,
  built_table,
  social_table
) %>%
  relocate(
    Domain,
    .before = Characteristic
  ) %>%
  pivot_wider(
    names_from = iso2,
    values_from = value
  ) %>%
  rename(
    Colombia = CO,
    `El Salvador` = SV,
    Guatemala = GT,
    Mexico = MX
  ) %>%
  relocate(
    Total,
    .after = Characteristic
  )

table1 %>%
  gt(
    groupname_col = "Domain"
  ) %>%
  cols_label(
    Characteristic = "",
    Total = "Total",
    Colombia = "Colombia",
    `El Salvador` = "El Salvador",
    Guatemala = "Guatemala",
    Mexico = "Mexico"
  ) %>%
  cols_align(
    align = "left",
    columns = Characteristic
  ) %>%
  cols_align(
    align = "center",
    columns = c(
      Total,
      Colombia,
      `El Salvador`,
      Guatemala,
      Mexico
    )
  ) %>%
  tab_header(
    title = md(
      "**Table 1. Characteristics of the sample by country**"
    )
  ) %>%
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    data_row.padding = px(3),
    row_group.font.weight = "bold"
  )
```