---
title: "Walking SALURBAL-Climate Proposal"
author:
  - name: Amanda Silva Magalhães
    affiliation: Federal University of Minas Gerais - Brazil
  - name: Amanda Cristina de Souza Andrade
    affiliation: René Rachou Institute, Oswaldo Cruz Foundation - Brazil 
  - name: Waleska Teixeira Caiaffa
    affiliation: Federal University of Minas Gerais - Brazil

date: today

bibliography: references.bib
---



# Setup

```{r}
#| label: Clean environment
#| code-summary: Clean environment
rm(list = ls())
```

```{r}
#| label: Directory
#| code-summary: Directory
setwd("C:/Users/amand/Amanda/GitHub/walking")
```

```{r}
#| label: Libraries
#| code-summary: Libraries
pacman::p_load(
  # Data import
  haven,
  readxl,
  # Data manipulation
  dplyr,
  tidyr,
  stringr,
  # Parallel processing
  parallel,
  pbapply,
  # Spatial data
  sf,
  # Statistical modeling
  splines,
  lme4,
  glmmTMB,
  mgcv,
  # Model predictions and diagnostics
  ggeffects,
  performance,
  DHARMa,
  # Tables and reporting
  gt,
  # Data visualization
  ggplot2,
  patchwork,
  corrplot
)
```

```{r}
#| label: Data
#| code-summary: Data
dta <- haven::read_dta("walking.dta")
```



# Distribution of outcomes


## Global walking

Global walking levels differed across countries. Median walking was higher in Mexico and El Salvador and lower in Colombia and Guatemala. Greater variability was observed in Mexico and El Salvador, while extreme values were present in all countries. Histogram frequencies represent the number of individuals interviewed within each walking duration (min/week) interval.

```{r}
#| label: Histogram global
#| code-summary: Histogram global
dta %>%
  mutate(
    iso2 = recode(
      iso2,
      CO = "Colombia",
      GT = "Guatemala",
      SV = "El Salvador",
      MX = "Mexico"
    )
  ) %>%
  ggplot(aes(x = svypawalk)) +
  geom_histogram(
    binwidth = 30,
    fill = "#4666A0",
    color = "white"
  ) +
  facet_wrap(~iso2, scales = "free_y") +
  labs(
    x = "\n Global walking (min/week)",
    y = "Frequency \n"
  ) +
  theme_minimal(base_size = 14)
```

```{r}
#| label: Boxplot global
#| code-summary: Boxplot global
dta %>%
  mutate(
    iso2 = recode(
      iso2,
      CO = "Colombia",
      GT = "Guatemala",
      SV = "El Salvador",
      MX = "Mexico"
    )
  ) %>%
  ggplot(aes(x = iso2, y = svypawalk, fill = iso2)) +
  geom_boxplot() +
  scale_fill_manual(
    values = c(
      "Colombia" = "#F0BF21",
      "Guatemala" = "#BCD5FF",
      "El Salvador" = "#303B56",
      "Mexico" = "#238179"
    )
  ) +
  scale_y_continuous(
    breaks = seq(0, 1500, by = 250)
  ) +
  labs(
    x = "",
    y = "Global walking (min/week) \n"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "none"
  )
```

## Leisure walking

Leisure walking levels were low in both countries, with median values close to zero. Greater variability and higher maximum values were observed in Colombia than in Guatemala. Histogram frequencies represent the number of individuals interviewed within each walking duration (min/week) interval.

```{r}
#| label: Histogram leisure
#| code-summary: Histogram leisure
dta %>%
  filter(iso2 %in% c("CO", "GT")) %>%
  mutate(
    iso2 = recode(
      iso2,
      CO = "Colombia",
      GT = "Guatemala"
    )
  ) %>%
  ggplot(aes(x = svypawlei)) +
  geom_histogram(
    binwidth = 30,
    fill = "#4666A0",
    color = "white"
  ) +
  facet_wrap(~iso2, scales = "free_y") +
  labs(
    x = "\n Leisure walking (min/week)",
    y = "Frequency \n"
  ) +
  theme_minimal(base_size = 14)
```

```{r}
#| label: Boxplot leisure
#| code-summary: Boxplot leisure
dta %>%
  filter(iso2 %in% c("CO", "GT")) %>%
  mutate(
    iso2 = recode(
      iso2,
      CO = "Colombia",
      GT = "Guatemala"
    )
  ) %>%
  ggplot(aes(x = iso2, y = svypawlei, fill = iso2)) +
  geom_boxplot() +
  scale_fill_manual(
    values = c(
      "Colombia" = "#F0BF21",
      "Guatemala" = "#BCD5FF"
    )
  ) +
  scale_y_continuous(
    breaks = seq(0, 1500, by = 250)
  ) +
  labs(
    x = "",
    y = "Leisure walking (min/week) \n"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "none"
  )
```

## Transportation walking

Transportation walking levels were higher in Colombia than in Guatemala, as indicated by a higher median value. Greater variability and higher maximum values were also observed in Colombia. Histogram frequencies represent the number of individuals interviewed within each walking duration (min/week) interval.

```{r}
#| label: Histogram transportation
#| code-summary: Histogram transportation
dta %>%
  filter(iso2 %in% c("CO", "GT")) %>%
  mutate(
    iso2 = recode(
      iso2,
      CO = "Colombia",
      GT = "Guatemala"
    )
  ) %>%
  ggplot(aes(x = svypawtrn)) +
  geom_histogram(
    binwidth = 30,
    fill = "#4666A0",
    color = "white"
  ) +
  facet_wrap(~iso2, scales = "free_y") +
  labs(
    x = "\n Transportation walking (min/week)",
    y = "Frequency \n"
  ) +
  theme_minimal(base_size = 14)
```

```{r}
#| label: Boxplot transportation
#| code-summary: Boxplot transportation
dta %>%
  filter(iso2 %in% c("CO", "GT")) %>%
  mutate(
    iso2 = recode(
      iso2,
      CO = "Colombia",
      GT = "Guatemala"
    )
  ) %>%
  ggplot(aes(x = iso2, y = svypawtrn, fill = iso2)) +
  geom_boxplot() +
  scale_fill_manual(
    values = c(
      "Colombia" = "#F0BF21",
      "Guatemala" = "#BCD5FF"
    )
  ) +
  scale_y_continuous(
    breaks = seq(0, 1500, by = 250)
  ) +
  labs(
    x = "",
    y = "Transportation walking (min/week) \n"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "none"
  )
```



# Distribution of exposure

The distribution of mean temperature during the week prior to the interview varied across countries. Guatemala and El Salvador showed relatively narrow temperature ranges, whereas Colombia and Mexico exhibited greater variability and a wider distribution of temperatures. Histogram frequencies represent the number of individuals interviewed within each temperature interval.

```{r}
#| label: Histogram temperature
#| code-summary: Histogram temperature
dta %>%
  mutate(
    iso2 = recode(
      iso2,
      CO = "Colombia",
      GT = "Guatemala",
      SV = "El Salvador",
      MX = "Mexico"
    )
  ) %>%
  ggplot(aes(x = l1adtemp_pw_1_7_mean)) +
  geom_histogram(
    binwidth = 1,
    fill = "#4666A0",
    color = "white"
  ) +
  facet_wrap(~iso2, scales = "free_y") +
  scale_x_continuous(
    breaks = seq(0, 35, by = 5)
  ) +
  labs(
    x = "\n Mean temperature (°C)",
    y = "Frequency \n"
  ) +
  theme_minimal(base_size = 14)
```

```{r}
#| label: Boxplot temperature
#| code-summary: Boxplot temperature
dta %>%
  mutate(
    iso2 = recode(
      iso2,
      CO = "Colombia",
      GT = "Guatemala",
      SV = "El Salvador",
      MX = "Mexico"
    )
  ) %>%
  ggplot(aes(x = iso2, y = l1adtemp_pw_1_7_mean, fill = iso2)) +
  geom_boxplot() +
  scale_fill_manual(
    values = c(
      "Colombia" = "#F0BF21",
      "Guatemala" = "#BCD5FF",
      "El Salvador" = "#303B56",
      "Mexico" = "#238179"
    )
  ) +
  scale_y_continuous(
    breaks = seq(0, 35, by = 5)
  ) +
  labs(
    x = "",
    y = "Mean temperature (°C) \n"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "none"
  )
```



# Outcome by exposure

All walking domains showed an apparent non-linear association with mean temperature during the week prior to the interview. These exploratory analyses were restricted to participants reporting more than 0 min/week of walking to focus on variation among individuals who engaged in walking activities.

```{r}
#| label: Histogram temperature by outcome
#| code-summary: Histogram temperature by outcome
temp_global <- dta %>%
  filter(
    !is.na(svypawalk),
    !is.na(l1adtemp_pw_1_7_mean)
  ) %>%
  transmute(
    outcome = "Global walking \n",
    temperature = l1adtemp_pw_1_7_mean
  )
temp_lei <- dta %>%
  filter(
    !is.na(svypawlei),
    !is.na(l1adtemp_pw_1_7_mean)
  ) %>%
  transmute(
    outcome = "Leisure walking \n",
    temperature = l1adtemp_pw_1_7_mean
  )
temp_trn <- dta %>%
  filter(
    !is.na(svypawtrn),
    !is.na(l1adtemp_pw_1_7_mean)
  ) %>%
  transmute(
    outcome = "Transportation walking \n",
    temperature = l1adtemp_pw_1_7_mean
  )
bind_rows(
  temp_global,
  temp_lei,
  temp_trn
) %>%
  ggplot(
    aes(x = temperature)
  ) +
  geom_histogram(
    binwidth = 1,
    fill = "#4666A0",
    color = "white"
  ) +
  facet_wrap(
    ~outcome,
    scales = "free_y"
  ) +
  scale_x_continuous(
    breaks = seq(0, 35, by = 5)
  ) +
  labs(
    x = "\n Mean temperature (°C)",
    y = "Frequency \n"
  ) +
  theme_minimal(base_size = 14)
```

```{r}
#| label: Global by temperature
#| code-summary: Global by temperature
dta %>%
  filter(svypawalk > 0) %>%
  ggplot(
    aes(
      x = l1adtemp_pw_1_7_mean,
      y = svypawalk
    )
  ) +
  geom_smooth(se = TRUE) +
  labs(
    x = "\n Mean temperature (°C)",
    y = "Global walking (min/week) \n"
  ) +
  theme_minimal(base_size = 14)
```

```{r}
#| label: Leisure by temperature
#| code-summary: Leisure by temperature
dta %>%
  filter(svypawlei > 0) %>%
  ggplot(
    aes(
      x = l1adtemp_pw_1_7_mean,
      y = svypawlei
    )
  ) +
  geom_smooth(se = TRUE) +
  labs(
    x = "\n Mean temperature (°C)",
    y = "Leisure walking (min/week) \n"
  ) +
  theme_minimal(base_size = 14)
```

```{r}
#| label: Transportation by temperature
#| code-summary: Transportation by temperature
dta %>%
  filter(svypawtrn > 0) %>%
  ggplot(
    aes(
      x = l1adtemp_pw_1_7_mean,
      y = svypawtrn
    )
  ) +
  geom_smooth(se = TRUE) +
  labs(
    x = "\n Mean temperature (°C)",
    y = "Transportation walking (min/week) \n"
  ) +
  theme_minimal(base_size = 14)
```



# Tabel 1

```{r}
#| label: Functions
#| code-summary: Functions
make_continuous_table <- function(data, variable, label, digits = 1) {
  safe_min <- function(x) {
    if (all(is.na(x))) NA_real_ else min(x, na.rm = TRUE)
  }
  safe_max <- function(x) {
    if (all(is.na(x))) NA_real_ else max(x, na.rm = TRUE)
  }
  total <- data.frame(
    iso2 = "Total",
    median = median(data[[variable]], na.rm = TRUE),
    Q1_val = quantile(data[[variable]], .25, na.rm = TRUE),
    Q3_val = quantile(data[[variable]], .75, na.rm = TRUE),
    min_val = safe_min(data[[variable]]),
    max_val = safe_max(data[[variable]])
  )
  by_iso2 <- data %>%
    filter(!is.na(iso2)) %>%
    group_by(iso2) %>%
    summarise(
      median = median(.data[[variable]], na.rm = TRUE),
      Q1_val = quantile(.data[[variable]], .25, na.rm = TRUE),
      Q3_val = quantile(.data[[variable]], .75, na.rm = TRUE),
      min_val = safe_min(.data[[variable]]),
      max_val = safe_max(.data[[variable]]),
      .groups = "drop"
    )
  stats <- bind_rows(total, by_iso2)
  bind_rows(
    stats %>%
      mutate(
        value = ifelse(
          is.na(median),
          NA_character_,
          paste0(
            format(round(median, digits), trim = TRUE),
            " (",
            format(round(Q1_val, digits), trim = TRUE),
            "; ",
            format(round(Q3_val, digits), trim = TRUE),
            ")"
          )
        ),
        Characteristic = paste0(label, ", median (Q1; Q3)")
      ) %>%
      select(Characteristic, iso2, value),
    stats %>%
      mutate(
        value = ifelse(
          is.na(min_val),
          NA_character_,
          paste0(
            format(round(min_val, digits), trim = TRUE),
            " - ",
            format(round(max_val, digits), trim = TRUE)
          )
        ),
        Characteristic = paste0(label, ", min-max")
      ) %>%
      select(Characteristic, iso2, value)
  )
}
make_categorical_table <- function(data, variable, label) {
  total <- data %>%
    filter(!is.na(.data[[variable]])) %>%
    count(.data[[variable]]) %>%
    mutate(
      iso2 = "Total",
      percent = 100 * n / sum(n),
      value = paste0(
        format(n, big.mark = ","),
        " (",
        formatC(percent, format = "f", digits = 1),
        ")"
      )
    ) %>%
    rename(level = all_of(variable)) %>%
    select(iso2, level, value)
  by_iso2 <- data %>%
    filter(!is.na(.data[[variable]]), !is.na(iso2)) %>%
    group_by(iso2, .data[[variable]]) %>%
    summarise(
      n = n(),
      .groups = "drop"
    ) %>%
    group_by(iso2) %>%
    mutate(
      percent = 100 * n / sum(n),
      value = paste0(
        format(n, big.mark = ","),
        " (",
        formatC(percent, format = "f", digits = 1),
        ")"
      )
    ) %>%
    ungroup() %>%
    rename(level = all_of(variable)) %>%
    select(iso2, level, value)
  bind_rows(total, by_iso2) %>%
    mutate(
      level = as.character(level),
      Characteristic = paste0("   ", label, " - ", level)
    ) %>%
    select(Characteristic, iso2, value) %>%
    bind_rows(
      tibble(
        Characteristic = label,
        iso2 = "Total",
        value = ""
      ),
      .
    )
}
```

```{r}
#| label: Context
#| code-summary: Context
# Number of cities
n_cities <- dta %>%
  filter(salid1_tag == 1) %>%
  count(iso2, name = "value") %>%
  bind_rows(
    summarise(., iso2 = "Total", value = sum(value))
  ) %>%
  mutate(
    value = format(
      value,
      big.mark = ",",
      decimal.mark = ".",
      scientific = FALSE
    ),
    Characteristic = "Number of cities"
  ) %>%
  select(Characteristic, iso2, value)
# Participants per city
participants_city <- make_continuous_table(
  data = dta %>% filter(!is.na(salid1)),
  variable = "salid1_obs",
  label = "Participants per city"
)
# Number of sub-cities
n_subcities <- dta %>%
  filter(salid2_tag == 1) %>%
  count(iso2, name = "value") %>%
  bind_rows(
    summarise(., iso2 = "Total", value = sum(value))
  ) %>%
  mutate(
    value = format(
      value,
      big.mark = ",",
      decimal.mark = "."
    ),
    Characteristic = "Number of sub-cities"
  ) %>%
  select(Characteristic, iso2, value)
# Participants per sub-city
participants_subcity <- make_continuous_table(
  data = dta %>% filter(!is.na(salid2)),
  variable = "salid2_obs",
  label = "Participants per sub-city"
)
# Number of participants
n_participants <- dta %>%
  count(iso2, name = "n") %>%
  bind_rows(
    summarise(., iso2 = "Total", n = sum(n))
  ) %>%
  mutate(
    pct = round(100 * n / n[iso2 == "Total"], 1),
    value = ifelse(
      iso2 == "Total",
      paste0(format(n, big.mark = ","), " (100.0)"),
      paste0(format(n, big.mark = ","), " (", pct, ")")
    ),
    Characteristic = "Number of participants, n (%)"
  ) %>%
  select(Characteristic, iso2, value)
# Context table
context_table <- bind_rows(
  n_cities,
  participants_city,
  n_subcities,
  participants_subcity,
  n_participants
)
context_table <- context_table %>%
  mutate(Domain = "Context")
```

```{r}
#| label: Individual-level
#| code-summary: Individual-level
# Recoding categorical variables
dta <- dta %>%
  mutate(
    svymale = factor(
      svymale,
      levels = c(0, 1),
      labels = c("Female", "Male")
    ),
    svyedu = case_when(
      svyedu %in% c(9998, 9999) ~ NA_real_,
      TRUE ~ as.numeric(svyedu)
    ),
    svyedu = factor(
      svyedu,
      levels = c(1, 2, 3, 4),
      labels = c(
        "Less than primary",
        "Primary",
        "Secondary",
        "University or higher"
      )
    ),
    svypawalkcat = factor(
      svypawalkcat,
      levels = c(0, 1),
      labels = c("Inactive", "Active")
    ),
    svypawleicat = factor(
      svypawleicat,
      levels = c(0, 1),
      labels = c("Inactive", "Active")
    ),
    svypawtrncat = factor(
      svypawtrncat,
      levels = c(0, 1),
      labels = c("Inactive", "Active")
    )
  )
# Age
age_total <- dta %>%
  summarise(
    iso2 = "Total",
    mean = mean(svyage, na.rm = TRUE),
    sd = sd(svyage, na.rm = TRUE),
    min_val = min(svyage, na.rm = TRUE),
    max_val = max(svyage, na.rm = TRUE)
  )
age_by_iso2 <- dta %>%
  group_by(iso2) %>%
  summarise(
    mean = mean(svyage, na.rm = TRUE),
    sd = sd(svyage, na.rm = TRUE),
    min_val = min(svyage, na.rm = TRUE),
    max_val = max(svyage, na.rm = TRUE),
    .groups = "drop"
  )
age_stats <- bind_rows(age_total, age_by_iso2)
age_table <- bind_rows(
  age_stats %>%
    mutate(
      value = paste0(
        format(round(mean, 1), trim = TRUE),
        " (",
        format(round(sd, 1), trim = TRUE),
        ")"
      ),
      Characteristic = "Age, mean (SD)"
    ) %>%
    select(Characteristic, iso2, value),
  age_stats %>%
    mutate(
      value = paste0(
        format(round(min_val, 1), trim = TRUE),
        " - ",
        format(round(max_val, 1), trim = TRUE)
      ),
      Characteristic = "Age, min-max"
    ) %>%
    select(Characteristic, iso2, value)
)
# Sex
sex_table <- make_categorical_table(
  dta,
  "svymale",
  "Sex, n (%)"
)
# Educational level
education_table <- make_categorical_table(
  dta,
  "svyedu",
  "Educational level, n (%)"
)
# Global walking
global_walk <- make_continuous_table(
  dta,
  "svypawalk",
  "Global walking (min/week)",
  digits = 0
)
global_walk_cat <- make_categorical_table(
  dta,
  "svypawalkcat",
  "Global walking, n (%)"
)
# Leisure walking
leisure_walk <- make_continuous_table(
  dta,
  "svypawlei",
  "Leisure walking (min/week)",
  digits = 0
)
leisure_walk_cat <- make_categorical_table(
  dta,
  "svypawleicat",
  "Leisure walking, n (%)"
)
# Transportation walking
transport_walk <- make_continuous_table(
  dta,
  "svypawtrn",
  "Transportation walking (min/week)",
  digits = 0
)
transport_walk_cat <- make_categorical_table(
  dta,
  "svypawtrncat",
  "Transportation walking, n (%)"
)
# Final individual-level table
individual_table <- bind_rows(
  age_table,
  sex_table,
  education_table,
  global_walk,
  global_walk_cat,
  leisure_walk,
  leisure_walk_cat,
  transport_walk,
  transport_walk_cat
) %>%
  mutate(
    Domain = "Individual-level"
  )
```

```{r}
#| label: Natural environment
#| code-summary: Natural environment
# Mean temperature
temperature_city <- dta %>%
  filter(!is.na(salid1)) %>%
  group_by(iso2, salid1) %>%
  summarise(
    value = mean(l1adtemp_pw_1_7_mean, na.rm = TRUE),
    .groups = "drop"
  )
temperature_total <- temperature_city %>%
  summarise(
    iso2 = "Total",
    median = median(value, na.rm = TRUE),
    P10 = quantile(value, .10, na.rm = TRUE),
    P90 = quantile(value, .90, na.rm = TRUE),
    min_val = min(value, na.rm = TRUE),
    max_val = max(value, na.rm = TRUE)
  )
temperature_by_iso2 <- temperature_city %>%
  group_by(iso2) %>%
  summarise(
    median = median(value, na.rm = TRUE),
    P10 = quantile(value, .10, na.rm = TRUE),
    P90 = quantile(value, .90, na.rm = TRUE),
    min_val = min(value, na.rm = TRUE),
    max_val = max(value, na.rm = TRUE),
    .groups = "drop"
  )
temperature_stats <- bind_rows(
  temperature_total,
  temperature_by_iso2
)
temperature_table <- bind_rows(
  temperature_stats %>%
    mutate(
      value = paste0(
        format(round(median, 1), trim = TRUE),
        " (",
        format(round(P10, 1), trim = TRUE),
        "; ",
        format(round(P90, 1), trim = TRUE),
        ")"
      ),
      Characteristic = "Mean temperature (°C), median (P10; P90)"
    ) %>%
    select(Characteristic, iso2, value),
  temperature_stats %>%
    mutate(
      value = paste0(
        format(round(min_val, 1), trim = TRUE),
        " - ",
        format(round(max_val, 1), trim = TRUE)
      ),
      Characteristic = "Mean temperature (°C), min-max"
    ) %>%
    select(Characteristic, iso2, value)
)
# Climate zone
dta <- dta %>%
  mutate(
    becczl1adcat = as.character(
      haven::as_factor(becczl1adcat)
    )
  )
climate_city <- dta %>%
  filter(
    !is.na(becczl1adcat),
    !is.na(salid1)
  ) %>%
  distinct(
    iso2,
    salid1,
    becczl1adcat
  )
climate_total <- climate_city %>%
  count(becczl1adcat) %>%
  mutate(
    iso2 = "Total",
    percent = 100 * n / sum(n),
    value = paste0(
      format(n, big.mark = ","),
      " (",
      formatC(percent, digits = 1, format = "f"),
      ")"
    )
  ) %>%
  transmute(
    Characteristic = paste0(
      "\u00A0\u00A0\u00A0",
      becczl1adcat
    ),
    iso2,
    value
  )
climate_by_iso2 <- climate_city %>%
  count(
    iso2,
    becczl1adcat
  ) %>%
  group_by(iso2) %>%
  mutate(
    percent = 100 * n / sum(n),
    value = paste0(
      format(n, big.mark = ","),
      " (",
      formatC(percent, digits = 1, format = "f"),
      ")"
    )
  ) %>%
  ungroup() %>%
  transmute(
    Characteristic = paste0(
      "\u00A0\u00A0\u00A0",
      becczl1adcat
    ),
    iso2,
    value
  )
climate_table <- bind_rows(
  tibble(
    Characteristic = "Major climate zone, n (%)",
    iso2 = "Total",
    value = ""
  ),
  climate_total,
  climate_by_iso2
)
# Greenness
greenness_subcity <- dta %>%
  filter(!is.na(salid2)) %>%
  group_by(
    iso2,
    salid2
  ) %>%
  summarise(
    value = mean(becmedndvinwl2, na.rm = TRUE),
    .groups = "drop"
  )
greenness_table <- make_continuous_table(
  greenness_subcity,
  "value",
  "Greenness",
  digits = 1
)
# Final natural environment table
natural_table <- bind_rows(
  temperature_table,
  climate_table,
  greenness_table
)
natural_table <- natural_table %>%
  mutate(Domain = "Natural environment")
```

```{r}
#| label: Built environment
#| code-summary: Built environment
# Presence of a BRT
brt_city <- dta %>%
  filter(
    !is.na(becprsbrtl1ad),
    !is.na(salid1)
  ) %>%
  distinct(
    iso2,
    salid1,
    becprsbrtl1ad
  ) %>%
  mutate(
    becprsbrtl1ad = factor(
      becprsbrtl1ad,
      levels = c(0, 1),
      labels = c("No", "Yes")
    )
  )
brt_total <- brt_city %>%
  count(becprsbrtl1ad) %>%
  mutate(
    iso2 = "Total",
    percent = 100 * n / sum(n),
    value = paste0(
      format(n, big.mark = ","),
      " (",
      formatC(percent, digits = 1, format = "f"),
      ")"
    )
  ) %>%
  transmute(
    Characteristic = paste0(
      "\u00A0\u00A0\u00A0",
      becprsbrtl1ad
    ),
    iso2,
    value
  )
brt_by_iso2 <- brt_city %>%
  count(
    iso2,
    becprsbrtl1ad
  ) %>%
  group_by(iso2) %>%
  mutate(
    percent = 100 * n / sum(n),
    value = paste0(
      format(n, big.mark = ","),
      " (",
      formatC(percent, digits = 1, format = "f"),
      ")"
    )
  ) %>%
  ungroup() %>%
  transmute(
    Characteristic = paste0(
      "\u00A0\u00A0\u00A0",
      becprsbrtl1ad
    ),
    iso2,
    value
  )
brt_table <- bind_rows(
  tibble(
    Characteristic = "Presence of a BRT, n (%)",
    iso2 = "Total",
    value = ""
  ),
  brt_total,
  brt_by_iso2
)
# Percentage of urban area
urban_area_city <- dta %>%
  filter(!is.na(salid1)) %>%
  group_by(
    iso2,
    salid1
  ) %>%
  summarise(
    value = mean(becpcturbanl1ad, na.rm = TRUE),
    .groups = "drop"
  )
urban_area_table <- make_continuous_table(
  urban_area_city,
  "value",
  "Percentage of urban area",
  digits = 1
)
# Population density
population_density_subcity <- dta %>%
  filter(!is.na(salid2)) %>%
  group_by(
    iso2,
    salid2
  ) %>%
  summarise(
    value = mean(becpopdensl2, na.rm = TRUE),
    .groups = "drop"
  )
population_density_table <- make_continuous_table(
  population_density_subcity,
  "value",
  "Population density",
  digits = 0
)
# Intersection density
intersection_density_subcity <- dta %>%
  filter(!is.na(salid2)) %>%
  group_by(
    iso2,
    salid2
  ) %>%
  summarise(
    value = mean(becadintdensl2, na.rm = TRUE),
    .groups = "drop"
  )
intersection_density_table <- make_continuous_table(
  intersection_density_subcity,
  "value",
  "Intersection density",
  digits = 1
)
# Parks density
parks_density_subcity <- dta %>%
  filter(!is.na(salid2)) %>%
  group_by(
    iso2,
    salid2
  ) %>%
  summarise(
    value = mean(becparknumpercapl2, na.rm = TRUE),
    .groups = "drop"
  )
parks_density_table <- make_continuous_table(
  parks_density_subcity,
  "value",
  "Parks density",
  digits = 1
)
# Final built environment table
built_table <- bind_rows(
  brt_table,
  urban_area_table,
  population_density_table,
  intersection_density_table,
  parks_density_table
)
built_table <- built_table %>%
  mutate(Domain = "Built environment")
```

```{r}
#| label: Social environment
#| code-summary: Social environment
# Social environment index
social_city <- dta %>%
  filter(!is.na(salid1)) %>%
  group_by(
    iso2,
    salid1
  ) %>%
  summarise(
    value = mean(cnssei_l1ad, na.rm = TRUE),
    .groups = "drop"
  )
social_table <- make_continuous_table(
  social_city,
  "value",
  "Social environment index",
  digits = 1
)
social_table <- social_table %>%
  mutate(Domain = "Social environment")
```

```{r}
#| label: Table 1
#| code-summary: Table 1
table1 <- bind_rows(
  context_table,
  individual_table,
  natural_table,
  built_table,
  social_table
) %>%
  relocate(
    Domain,
    .before = Characteristic
  ) %>%
  pivot_wider(
    names_from = iso2,
    values_from = value
  ) %>%
  rename(
    Colombia = CO,
    `El Salvador` = SV,
    Guatemala = GT,
    Mexico = MX
  ) %>%
  relocate(
    Total,
    .after = Characteristic
  ) %>%
  mutate(
    is_main_var = !str_detect(Characteristic, " - "),
    across(
      c(
        Total,
        Colombia,
        Guatemala,
        Mexico,
        `El Salvador`
      ),
      ~ replace_na(.x, "")
    ),
    Characteristic = case_when(
      str_detect(Characteristic, " - ") ~
        paste0(
          "\u00A0\u00A0\u00A0",
          str_remove(Characteristic, "^.* - ")
        ),
      TRUE ~ Characteristic
    )
  )
table1 %>%
  gt(
    groupname_col = "Domain"
  ) %>%
  cols_hide(
    columns = is_main_var
  ) %>%
  cols_label(
    Characteristic = "",
    Total = "Total",
    Colombia = "Colombia",
    Guatemala = "Guatemala",
    Mexico = "Mexico",
    `El Salvador` = "El Salvador"
  ) %>%
  cols_align(
    align = "left",
    columns = Characteristic
  ) %>%
  cols_align(
    align = "center",
    columns = c(
      Total,
      Colombia,
      Guatemala,
      Mexico,
      `El Salvador`
    )
  ) %>%
  tab_header(
    title = md(
      "**Table 1. Characteristics of the sample by country**"
    )
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels(everything())
  ) %>%
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    data_row.padding = px(3),
    row_group.font.weight = "bold"
  )
```



# Knots

```{r}
#| label: Knots overall
#| code-summary: Knots overall
#| results: hide
# Define candidate knot locations
k1_vals <- c(0.10, 0.15, 0.25)
k2_vals <- c(0.50, 0.60, 0.75)
k3_vals <- c(0.85, 0.90, 0.95)
knot_options <- list()
for (k1 in k1_vals) {
  for (k2 in k2_vals) {
    for (k3 in k3_vals) {
      if (k1 < k2 & k2 < k3) {
        knot_options[[length(knot_options) + 1]] <- c(k1, k2, k3)
      }
    }
  }
}
names(knot_options) <- sapply(knot_options, function(x) {
  paste0(
    "k_",
    sprintf("%02d", round(x[1] * 100)),
    "_",
    sprintf("%02d", round(x[2] * 100)),
    "_",
    sprintf("%02d", round(x[3] * 100))
  )
})
# Global walking
pred_knots_global <- quantile(
  dta$l1adtemp_pw_1_7_mean[
    !is.na(dta$svypawalk) &
      !is.na(dta$l1adtemp_pw_1_7_mean)
  ],
  probs = c(0.25, 0.60, 0.85),
  na.rm = TRUE
)
pred_knots_global
# Leisure walking
pred_knots_lei <- quantile(
  dta$l1adtemp_pw_1_7_mean[
    !is.na(dta$svypawlei) &
      !is.na(dta$l1adtemp_pw_1_7_mean)
  ],
  probs = c(0.25, 0.75, 0.95),
  na.rm = TRUE
)
pred_knots_lei
# Transportation walking
pred_knots_trn <- quantile(
  dta$l1adtemp_pw_1_7_mean[
    !is.na(dta$svypawtrn) &
      !is.na(dta$l1adtemp_pw_1_7_mean)
  ],
  probs = c(0.10, 0.50, 0.85),
  na.rm = TRUE
)
pred_knots_trn
```

```{r}
#| label: Knots by country
#| code-summary: Knots by country
#| results: hide
countries <- sort(unique(na.omit(dta$iso2)))

# Global walking
pred_knots_global_country <- lapply(countries, function(ctry) {
  quantile(
    dta$l1adtemp_pw_1_7_mean[
      dta$iso2 == ctry &
        !is.na(dta$svypawalk) &
        !is.na(dta$l1adtemp_pw_1_7_mean)
    ],
    probs = c(0.25, 0.60, 0.85),
    na.rm = TRUE
  )
})
names(pred_knots_global_country) <- countries
pred_knots_global_country

# Leisure walking
pred_knots_lei_country <- lapply(countries, function(ctry) {
  quantile(
    dta$l1adtemp_pw_1_7_mean[
      dta$iso2 == ctry &
        !is.na(dta$svypawlei) &
        !is.na(dta$l1adtemp_pw_1_7_mean)
    ],
    probs = c(0.25, 0.75, 0.95),
    na.rm = TRUE
  )
})
names(pred_knots_lei_country) <- countries
pred_knots_lei_country
# Transportation walking
pred_knots_trn_country <- lapply(countries, function(ctry) {
  quantile(
    dta$l1adtemp_pw_1_7_mean[
      dta$iso2 == ctry &
        !is.na(dta$svypawtrn) &
        !is.na(dta$l1adtemp_pw_1_7_mean)
    ],
    probs = c(0.10, 0.50, 0.85),
    na.rm = TRUE
  )
})
names(pred_knots_trn_country) <- countries
pred_knots_trn_country
```



# Center year and standardize continuous covariates

```{r}
#| label: Center year and standardize continuous covariates
#| code-summary: Center year and standardize continuous covariates
#| results: hide
# Center year so that the first study year equals zero
dta$year_c <- dta$year - min(dta$year, na.rm = TRUE)
unique_years <- unique(
  st_drop_geometry(dta)[, c("year", "year_c")]
)
unique_years <- unique_years[
  order(unique_years$year),
]
print(unique_years)
# Convert quarter variable to a factor for modeling
dta <- dta %>%
  mutate(
    quarters = factor(quarters)
  )
## Standardize continuous covariates (mean = 0, SD = 1)
vars_to_scale <- c(
  "svyage",
  "becmedndvinwl2",
  "becpcturbanl1ad",
  "becpopdensl2",
  "becadintdensl2",
  "becparknumpercapl2",
  "cnssei_l1ad"
)
dta[paste0(vars_to_scale, "_scaled")] <- lapply(
  dta[vars_to_scale],
  function(x) as.numeric(scale(x))
)
```



# Model selection

Compare Poisson, negative binomial, and zero-inflated models

## Global walking

```{r}
#| label: Model selection global
#| code-summary: Model selection global
# # Poisson
# M0_global_pois <- glmmTMB(
#   svypawalk ~
#     ns(
#       l1adtemp_pw_1_7_mean,
#       knots = pred_knots_global
#     ) +
#     (1 | salid1),
#   data = dta,
#   family = poisson
# )
# # Negative binomial 1
# M0_global_nb1 <- glmmTMB(
#   svypawalk ~
#     ns(
#       l1adtemp_pw_1_7_mean,
#       knots = pred_knots_global
#     ) +
#     (1 | salid1),
#   data = dta,
#   family = nbinom1
# )
# # Negative binomial 2
# M0_global_nb2 <- glmmTMB(
#   svypawalk ~
#     ns(
#       l1adtemp_pw_1_7_mean,
#       knots = pred_knots_global
#     ) +
#     (1 | salid1),
#   data = dta,
#   family = nbinom2
# )
# # Zero-inflated Poisson
# M0_global_zip <- glmmTMB(
#   svypawalk ~
#     ns(
#       l1adtemp_pw_1_7_mean,
#       knots = pred_knots_global
#     ) +
#     (1 | salid1),
#   ziformula = ~1,
#   data = dta,
#   family = poisson
# )
# # Zero-inflated Negative Binomial 1
# M0_global_zinb1 <- glmmTMB(
#   svypawalk ~
#     ns(
#       l1adtemp_pw_1_7_mean,
#       knots = pred_knots_global
#     ) +
#     (1 | salid1),
#   ziformula = ~1,
#   data = dta,
#   family = nbinom1
# )
# # Zero-inflated Negative Binomial 2
# M0_global_zinb2 <- glmmTMB(
#   svypawalk ~
#     ns(
#       l1adtemp_pw_1_7_mean,
#       knots = pred_knots_global
#     ) +
#     (1 | salid1),
#   ziformula = ~1,
#   data = dta,
#   family = nbinom2
# )
# # Compare model fit
# AIC(
#   M0_global_pois,
#   M0_global_nb1,
#   M0_global_nb2,
#   M0_global_zip,
#   M0_global_zinb1,
#   M0_global_zinb2
# )
```

## Leisure walking

```{r}
#| label: Model selection leisure
#| code-summary: Model selection leisure
# # Poisson
# M0_lei_pois <- glmmTMB(
#   svypawlei ~
#     ns(
#       l1adtemp_pw_1_7_mean,
#       knots = pred_knots_domain
#     ) +
#     (1 | salid1),
#   data = dta,
#   family = poisson
# )
# # Negative binomial 1
# M0_lei_nb1 <- glmmTMB(
#   svypawlei ~
#     ns(
#       l1adtemp_pw_1_7_mean,
#       knots = pred_knots_domain
#     ) +
#     (1 | salid1),
#   data = dta,
#   family = nbinom1
# )
# # Negative binomial 2
# M0_lei_nb2 <- glmmTMB(
#   svypawlei ~
#     ns(
#       l1adtemp_pw_1_7_mean,
#       knots = pred_knots_domain
#     ) +
#     (1 | salid1),
#   data = dta,
#   family = nbinom2
# )
# # Zero-inflated Poisson
# M0_lei_zip <- glmmTMB(
#   svypawlei ~
#     ns(
#       l1adtemp_pw_1_7_mean,
#       knots = pred_knots_domain
#     ) +
#     (1 | salid1),
#   ziformula = ~1,
#   data = dta,
#   family = poisson
# )
# # Zero-inflated Negative Binomial 1
# M0_lei_zinb1 <- glmmTMB(
#   svypawlei ~
#     ns(
#       l1adtemp_pw_1_7_mean,
#       knots = pred_knots_domain
#     ) +
#     (1 | salid1),
#   ziformula = ~1,
#   data = dta,
#   family = nbinom1
# )
# # Zero-inflated Negative Binomial 2
# M0_lei_zinb2 <- glmmTMB(
#   svypawlei ~
#     ns(
#       l1adtemp_pw_1_7_mean,
#       knots = pred_knots_domain
#     ) +
#     (1 | salid1),
#   ziformula = ~1,
#   data = dta,
#   family = nbinom2
# )
# # Compare model fit
# AIC(
#   M0_lei_pois,
#   M0_lei_nb1,
#   M0_lei_nb2,
#   M0_lei_zip,
#   M0_lei_zinb1,
#   M0_lei_zinb2
# )
```

## Transportation walking

```{r}
#| label: Model selection transportation
#| code-summary: Model selection transportation
# # Poisson
# M0_trn_pois <- glmmTMB(
#   svypawtrn ~
#     ns(
#       l1adtemp_pw_1_7_mean,
#       knots = pred_knots_domain
#     ) +
#     (1 | salid1),
#   data = dta,
#   family = poisson
# )
# # Negative binomial 1
# M0_trn_nb1 <- glmmTMB(
#   svypawtrn ~
#     ns(
#       l1adtemp_pw_1_7_mean,
#       knots = pred_knots_domain
#     ) +
#     (1 | salid1),
#   data = dta,
#   family = nbinom1
# )
# # Negative binomial 2
# M0_trn_nb2 <- glmmTMB(
#   svypawtrn ~
#     ns(
#       l1adtemp_pw_1_7_mean,
#       knots = pred_knots_domain
#     ) +
#     (1 | salid1),
#   data = dta,
#   family = nbinom2
# )
# # Zero-inflated Poisson
# M0_trn_zip <- glmmTMB(
#   svypawtrn ~
#     ns(
#       l1adtemp_pw_1_7_mean,
#       knots = pred_knots_domain
#     ) +
#     (1 | salid1),
#   ziformula = ~1,
#   data = dta,
#   family = poisson
# )
# # Zero-inflated Negative Binomial 1
# M0_trn_zinb1 <- glmmTMB(
#   svypawtrn ~
#     ns(
#       l1adtemp_pw_1_7_mean,
#       knots = pred_knots_domain
#     ) +
#     (1 | salid1),
#   ziformula = ~1,
#   data = dta,
#   family = nbinom1
# )
# # Zero-inflated Negative Binomial 2
# M0_trn_zinb2 <- glmmTMB(
#   svypawtrn ~
#     ns(
#       l1adtemp_pw_1_7_mean,
#       knots = pred_knots_domain
#     ) +
#     (1 | salid1),
#   ziformula = ~1,
#   data = dta,
#   family = nbinom2
# )
# # Compare model fit
# AIC(
#   M0_trn_pois,
#   M0_trn_nb1,
#   M0_trn_nb2,
#   M0_trn_zip,
#   M0_trn_zinb1,
#   M0_trn_zinb2
# )
```



# Knot sensitivity analysis

```{r}
#| label: Function
#| code-summary: Function
run_knot_sensitivity <- function(
  outcome_var,
  outcome_label,
  knot_options,
  dta,
  cl
) {
  covars <- c(
    "svyage_scaled",
    "svymale",
    "svyedu",
    "becmedndvinwl2_scaled",
    "becprsbrtl1ad",
    "becpopdensl2_scaled",
    "becadintdensl2_scaled",
    "becparknumpercapl2_scaled",
    "cnssei_l1ad_scaled"
  )
  models <- pblapply(
    seq_along(knot_options),
    function(i) {
      tryCatch({
        valid_idx <- complete.cases(
          dta[, c(
            outcome_var,
            "l1adtemp_pw_1_7_mean",
            "salid1",
            "iso2",
            "quarters",
            "year_c",
            covars
          )]
        )
        knots <- quantile(
          dta$l1adtemp_pw_1_7_mean[valid_idx],
          probs = knot_options[[i]],
          na.rm = TRUE
        )
        formula_model <- as.formula(
          paste0(
            outcome_var,
            " ~ ",
            "splines::ns(l1adtemp_pw_1_7_mean, knots = knots) + ",
            "(1 | salid1) + ",
            "factor(iso2) + ",
            "factor(quarters) + ",
            "year_c + ",
            paste(covars, collapse = " + ")
          )
        )
        glmmTMB(
          formula_model,
          ziformula = ~1,
          data = dta,
          family = nbinom2
        )
      }, error = function(e) {
        message(
          paste(
            "Model error:",
            outcome_label,
            "-",
            names(knot_options)[i],
            "-",
            e$message
          )
        )
        NULL
      })
    },
    cl = cl
  )
  names(models) <- names(knot_options)
  results <- data.frame(
    outcome = outcome_label,
    model = names(models),
    AIC = sapply(
      models,
      function(m) {
        if (is.null(m)) NA else AIC(m)
      }
    )
  ) |>
    dplyr::filter(!is.na(AIC)) |>
    dplyr::arrange(AIC) |>
    dplyr::mutate(
      delta_AIC = AIC - min(AIC)
    )
  return(results)
}
```

```{r}
#| label: Parallel setup
#| code-summary: Parallel setup
n_cores <- max(1, detectCores() - 2)
cl <- makeCluster(n_cores)
clusterExport(
  cl,
  varlist = c(
    "dta",
    "knot_options"
  ),
  envir = environment()
)
clusterEvalQ(cl, {
  library(glmmTMB)
  library(splines)
})
```

```{r}
#| label: Knots global
#| code-summary: Knots global
# results_global <- run_knot_sensitivity(
#   outcome_var = "svypawalk",
#   outcome_label = "Global walking",
#   knot_options = knot_options,
#   dta = dta,
#   cl = cl
# )
```

```{r}
#| label: Knots leisure
#| code-summary: Knots leisure
# results_lei <- run_knot_sensitivity(
#   outcome_var = "svypawlei",
#   outcome_label = "Leisure walking",
#   knot_options = knot_options,
#   dta = dta,
#   cl = cl
# )
```

```{r}
#| label: Knots transportation
#| code-summary: Knots transportation
# results_trn <- run_knot_sensitivity(
#   outcome_var = "svypawtrn",
#   outcome_label = "Transportation walking",
#   knot_options = knot_options,
#   dta = dta,
#   cl = cl
# )
```

```{r}
#| label: Knots output
#| code-summary: Knots output
# # Stop cluster
# stopCluster(cl)
# # Combine results
# all_results <- bind_rows(
#   results_global,
#   results_lei,
#   results_trn
# ) %>%
#   arrange(
#     outcome,
#     delta_AIC
#   )
# # Save results
# write.table(
#   all_results,
#   file = "knot_sensitivity_results.txt",
#   sep = "\t",
#   row.names = FALSE,
#   quote = FALSE
# )
```



# Base models

## Global walking

```{r}
#| label: Base model global with penalized spline
#| code-summary: Base model global with penalized spline
M1_global_pspline <- glmmTMB(
  svypawalk ~
    s(
      l1adtemp_pw_1_7_mean,
      bs = "ps",
      k = 10
    ) +
    (1 | salid1) +
    factor(iso2) +
    factor(quarters) +
    year_c,
  ziformula = ~1,
  data = dta,
  family = nbinom2
)
```

```{r}
#| label: Base model global
#| code-summary: Base model global
M1_global_zinb2 <- glmmTMB(
  svypawalk ~
    splines::ns(l1adtemp_pw_1_7_mean, knots = pred_knots_global) +
    (1 | salid1) +
    factor(iso2) +
    factor(quarters) +
    year_c,
  ziformula = ~1,
  data = dta,
  family = nbinom2
)
summary(M1_global_zinb2)
```

```{r}
#| label: AIC comparison global models
#| code-summary: AIC comparison of global models
AIC(M1_global_pspline, M1_global_zinb2)
```

```{r}
#| label: IRR table global
#| code-summary: IRR table global
# Generate predictions across the full temperature range
pred_global <- ggpredict(
  M1_global_zinb2,
  terms = "l1adtemp_pw_1_7_mean [all]",
  bias_correction = TRUE
)
# Convert to data frame
pred_global_df <- as.data.frame(pred_global) %>%
  rename(
    temperature = x,
    predicted = predicted,
    lower = conf.low,
    upper = conf.high
  )
# Define reference temperature (25th percentile)
ref_temp <- pred_knots_global[[1]]
ref_index <- which.min(
  abs(pred_global_df$temperature - ref_temp)
)
ref_rate <- pred_global_df$predicted[ref_index]
# Calculate IRR
pred_global_df <- pred_global_df %>%
  mutate(
    IRR = predicted / ref_rate,
    IRR_low = lower / ref_rate,
    IRR_high = upper / ref_rate
  )
# Extract estimates at key percentiles
temps_interest <- as.numeric(pred_knots_global)
irr_global_table <- lapply(
  temps_interest,
  function(temp) {
    pred_global_df %>%
      slice(
        which.min(abs(temperature - temp))
      )
  }
) %>%
  bind_rows() %>%
  mutate(
    percentile = c(
      "25th percentile",
      "60th percentile",
      "85th percentile"
    ),
    IRR_CI = paste0(
      round(IRR, 2),
      " (",
      round(IRR_low, 2),
      "; ",
      round(IRR_high, 2),
      ")"
    )
  ) %>%
  select(
    percentile,
    temperature,
    IRR_CI
  )
irr_global_table
```

```{r}
#| label: Curve global
#| code-summary: Curve global
ggplot(
  pred_global_df,
  aes(
    x = temperature,
    y = IRR
  )
) +
  geom_ribbon(
    aes(
      ymin = IRR_low,
      ymax = IRR_high
    ),
    alpha = 0.20
  ) +
  geom_line(
    linewidth = 1
  ) +
  geom_hline(
    yintercept = 1,
    linetype = "dashed"
  ) +
  geom_vline(
    xintercept = ref_temp,
    linetype = "dashed"
  ) +
  labs(
    x = "Mean temperature (°C)",
    y = "Incidence Rate Ratio (IRR)"
  ) +
  theme_minimal()
```

## Leisure walking

```{r}
#| label: Base model leisure with penalized spline
#| code-summary: Base model leisure with penalized spline
M1_lei_pspline <- glmmTMB(
  svypawlei ~
    s(
      l1adtemp_pw_1_7_mean,
      bs = "ps",
      k = 10
    ) +
    (1 | salid1) +
    factor(iso2) +
    factor(quarters) +
    year_c,
  ziformula = ~1,
  data = dta,
  family = nbinom2
)
```

```{r}
#| label: Base model leisure
#| code-summary: Base model leisure
M1_lei_zinb2 <- glmmTMB(
  svypawlei ~
    splines::ns(l1adtemp_pw_1_7_mean, knots = pred_knots_lei) +
    (1 | salid1) +
    factor(iso2) +
    factor(quarters) +
    year_c,
  ziformula = ~1,
  data = dta,
  family = nbinom2
)
summary(M1_lei_zinb2)
```

```{r}
#| label: AIC comparison leisure models
#| code-summary: AIC comparison of leisure models
AIC(M1_lei_pspline, M1_lei_zinb2)
```

```{r}
#| label: IRR table leisure
#| code-summary: IRR table leisure
pred_lei <- ggpredict(
  M1_lei_zinb2,
  terms = "l1adtemp_pw_1_7_mean [all]",
  bias_correction = TRUE
)
pred_lei_df <- as.data.frame(pred_lei) %>%
  rename(
    temperature = x,
    predicted = predicted,
    lower = conf.low,
    upper = conf.high
  )
ref_temp <- pred_knots_lei[[1]]
ref_index <- which.min(
  abs(pred_lei_df$temperature - ref_temp)
)
ref_rate <- pred_lei_df$predicted[ref_index]
pred_lei_df <- pred_lei_df %>%
  mutate(
    IRR = predicted / ref_rate,
    IRR_low = lower / ref_rate,
    IRR_high = upper / ref_rate
  )
temps_interest <- as.numeric(pred_knots_lei)
irr_lei_table <- lapply(
  temps_interest,
  function(temp) {
    pred_lei_df %>%
      slice(
        which.min(abs(temperature - temp))
      )
  }
) %>%
  bind_rows() %>%
  mutate(
    percentile = c(
      "25th percentile",
      "75th percentile",
      "95th percentile"
    ),
    IRR_CI = paste0(
      round(IRR, 2),
      " (",
      round(IRR_low, 2),
      "; ",
      round(IRR_high, 2),
      ")"
    )
  ) %>%
  select(
    percentile,
    temperature,
    IRR_CI
  )
irr_lei_table
```

```{r}
#| label: Curve leisure
#| code-summary: Curve leisure
ggplot(
  pred_lei_df,
  aes(
    x = temperature,
    y = IRR
  )
) +
  geom_ribbon(
    aes(
      ymin = IRR_low,
      ymax = IRR_high
    ),
    alpha = 0.20
  ) +
  geom_line(
    linewidth = 1
  ) +
  geom_hline(
    yintercept = 1,
    linetype = "dashed"
  ) +
  geom_vline(
    xintercept = ref_temp,
    linetype = "dashed"
  ) +
  labs(
    x = "Mean temperature (°C)",
    y = "Incidence Rate Ratio (IRR)"
  ) +
  theme_minimal()
```

## Transportation walking

```{r}
#| label: Base model transportation with penalized spline
#| code-summary: Base model transportation with penalized spline
M1_trn_pspline <- glmmTMB(
  svypawtrn ~
    s(
      l1adtemp_pw_1_7_mean,
      bs = "ps",
      k = 10
    ) +
    (1 | salid1) +
    factor(iso2) +
    factor(quarters) +
    year_c,
  ziformula = ~1,
  data = dta,
  family = nbinom2
)
```

```{r}
#| label: Base model transportation
#| code-summary: Base model transportation
M1_trn_zinb2 <- glmmTMB(
  svypawtrn ~
    splines::ns(l1adtemp_pw_1_7_mean, knots = pred_knots_trn) +
    (1 | salid1) +
    factor(iso2) +
    factor(quarters) +
    year_c,
  ziformula = ~1,
  data = dta,
  family = nbinom2
)
summary(M1_trn_zinb2)
```

```{r}
#| label: AIC comparison transportation models
#| code-summary: AIC comparison of transportation models
AIC(M1_trn_pspline, M1_trn_zinb2)
```

```{r}
#| label: IRR table transportation
#| code-summary: IRR table transportation
pred_trn <- ggpredict(
  M1_trn_zinb2,
  terms = "l1adtemp_pw_1_7_mean [all]",
  bias_correction = TRUE
)
pred_trn_df <- as.data.frame(pred_trn) %>%
  rename(
    temperature = x,
    predicted = predicted,
    lower = conf.low,
    upper = conf.high
  )
ref_temp <- pred_knots_trn[[1]]
ref_index <- which.min(
  abs(pred_trn_df$temperature - ref_temp)
)
ref_rate <- pred_trn_df$predicted[ref_index]
pred_trn_df <- pred_trn_df %>%
  mutate(
    IRR = predicted / ref_rate,
    IRR_low = lower / ref_rate,
    IRR_high = upper / ref_rate
  )
temps_interest <- as.numeric(pred_knots_trn)
irr_trn_table <- lapply(
  temps_interest,
  function(temp) {
    pred_trn_df %>%
      slice(
        which.min(abs(temperature - temp))
      )
  }
) %>%
  bind_rows() %>%
  mutate(
    percentile = c(
      "10th percentile",
      "50th percentile",
      "85th percentile"
    ),
    IRR_CI = paste0(
      round(IRR, 2),
      " (",
      round(IRR_low, 2),
      "; ",
      round(IRR_high, 2),
      ")"
    )
  ) %>%
  select(
    percentile,
    temperature,
    IRR_CI
  )
irr_trn_table
```

```{r}
#| label: Curve transportation
#| code-summary: Curve transportation
ggplot(
  pred_trn_df,
  aes(
    x = temperature,
    y = IRR
  )
) +
  geom_ribbon(
    aes(
      ymin = IRR_low,
      ymax = IRR_high
    ),
    alpha = 0.20
  ) +
  geom_line(
    linewidth = 1
  ) +
  geom_hline(
    yintercept = 1,
    linetype = "dashed"
  ) +
  geom_vline(
    xintercept = ref_temp,
    linetype = "dashed"
  ) +
  labs(
    x = "Mean temperature (°C)",
    y = "Incidence Rate Ratio (IRR)"
  ) +
  theme_minimal()
```



# Fully adjusted models

## Global walking

```{r}
#| label: Full adjusted model global
#| code-summary: Full adjusted model global
M2_global_zinb2 <- glmmTMB(
  svypawalk ~
    splines::ns(l1adtemp_pw_1_7_mean, knots = pred_knots_global) +
    (1 | salid1) +
    factor(iso2) +
    factor(quarters) +
    year_c +
    svyage_scaled + svymale + svyedu +
    becmedndvinwl2_scaled +
    becprsbrtl1ad + becpopdensl2_scaled + 
    becadintdensl2_scaled + becparknumpercapl2_scaled +
    cnssei_l1ad_scaled,
  ziformula = ~1,
  data = dta,
  family = nbinom2
)
summary(M2_global_zinb2)
check_collinearity(M2_global_zinb2)
```

```{r}
#| label: IRR table global (full model)
#| code-summary: IRR table global (full model)
pred_global <- ggpredict(
  M2_global_zinb2,
  terms = "l1adtemp_pw_1_7_mean [all]",
  bias_correction = TRUE
)
pred_global_df <- as.data.frame(pred_global) %>%
  rename(
    temperature = x,
    predicted = predicted,
    lower = conf.low,
    upper = conf.high
  )
ref_temp <- pred_knots_global[[1]]
ref_index <- which.min(
  abs(pred_global_df$temperature - ref_temp)
)
ref_rate <- pred_global_df$predicted[ref_index]
pred_global_df <- pred_global_df %>%
  mutate(
    IRR = predicted / ref_rate,
    IRR_low = lower / ref_rate,
    IRR_high = upper / ref_rate
  )
temps_interest <- as.numeric(pred_knots_global)
irr_global_table <- lapply(
  temps_interest,
  function(temp) {
    pred_global_df %>%
      slice(
        which.min(abs(temperature - temp))
      )
  }
) %>%
  bind_rows() %>%
  mutate(
    percentile = c(
      "25th percentile",
      "60th percentile",
      "85th percentile"
    ),
    IRR_CI = paste0(
      round(IRR, 2),
      " (",
      round(IRR_low, 2),
      "; ",
      round(IRR_high, 2),
      ")"
    )
  ) %>%
  select(
    percentile,
    temperature,
    IRR_CI
  )
irr_global_table
```

```{r}
#| label: Curve global (full model)
#| code-summary: Curve global (full model)
ggplot(
  pred_global_df,
  aes(
    x = temperature,
    y = IRR
  )
) +
  geom_ribbon(
    aes(
      ymin = IRR_low,
      ymax = IRR_high
    ),
    alpha = 0.20
  ) +
  geom_line(
    linewidth = 1
  ) +
  geom_hline(
    yintercept = 1,
    linetype = "dashed"
  ) +
  geom_vline(
    xintercept = ref_temp,
    linetype = "dashed"
  ) +
  labs(
    x = "Mean temperature (°C)",
    y = "Incidence Rate Ratio (IRR)"
  ) +
  theme_minimal()
```

## Leisure walking

```{r}
#| label: Full adjusted model leisure
#| code-summary: Full adjusted model leisure
M2_lei_zinb2 <- glmmTMB(
  svypawlei ~
    splines::ns(l1adtemp_pw_1_7_mean, knots = pred_knots_lei) +
    (1 | salid1) +
    factor(iso2) +
    factor(quarters) +
    year_c +
    svyage_scaled + svymale + svyedu +
    becmedndvinwl2_scaled +
    becprsbrtl1ad + becpopdensl2_scaled + 
    becadintdensl2_scaled + becparknumpercapl2_scaled +
    cnssei_l1ad_scaled,
  ziformula = ~1,
  data = dta,
  family = nbinom2
)
summary(M2_lei_zinb2)
check_collinearity(M2_lei_zinb2)
```

```{r}
#| label: IRR table leisure (full model)
#| code-summary: IRR table leisure (full model)
pred_lei <- ggpredict(
  M2_lei_zinb2,
  terms = "l1adtemp_pw_1_7_mean [all]",
  bias_correction = TRUE
)
pred_lei_df <- as.data.frame(pred_lei) %>%
  rename(
    temperature = x,
    predicted = predicted,
    lower = conf.low,
    upper = conf.high
  )
ref_temp <- pred_knots_lei[[1]]
ref_index <- which.min(
  abs(pred_lei_df$temperature - ref_temp)
)
ref_rate <- pred_lei_df$predicted[ref_index]
pred_lei_df <- pred_lei_df %>%
  mutate(
    IRR = predicted / ref_rate,
    IRR_low = lower / ref_rate,
    IRR_high = upper / ref_rate
  )
temps_interest <- as.numeric(pred_knots_lei)
irr_lei_table <- lapply(
  temps_interest,
  function(temp) {
    pred_lei_df %>%
      slice(
        which.min(abs(temperature - temp))
      )
  }
) %>%
  bind_rows() %>%
  mutate(
    percentile = c(
      "25th percentile",
      "75th percentile",
      "95th percentile"
    ),
    IRR_CI = paste0(
      round(IRR, 2),
      " (",
      round(IRR_low, 2),
      "; ",
      round(IRR_high, 2),
      ")"
    )
  ) %>%
  select(
    percentile,
    temperature,
    IRR_CI
  )
irr_lei_table
```

```{r}
#| label: Curve leisure (full model)
#| code-summary: Curve leisure (full model)
ggplot(
  pred_lei_df,
  aes(
    x = temperature,
    y = IRR
  )
) +
  geom_ribbon(
    aes(
      ymin = IRR_low,
      ymax = IRR_high
    ),
    alpha = 0.20
  ) +
  geom_line(
    linewidth = 1
  ) +
  geom_hline(
    yintercept = 1,
    linetype = "dashed"
  ) +
  geom_vline(
    xintercept = ref_temp,
    linetype = "dashed"
  ) +
  labs(
    x = "Mean temperature (°C)",
    y = "Incidence Rate Ratio (IRR)"
  ) +
  theme_minimal()
```

## Transportation walking

```{r}
#| label: Full adjusted model transportation
#| code-summary: Full adjusted model transportation
M2_trn_zinb2 <- glmmTMB(
  svypawtrn ~
    splines::ns(l1adtemp_pw_1_7_mean, knots = pred_knots_trn) +
    (1 | salid1) +
    factor(iso2) +
    factor(quarters) +
    year_c +
    svyage_scaled + svymale + svyedu +
    becmedndvinwl2_scaled +
    becprsbrtl1ad + becpopdensl2_scaled + 
    becadintdensl2_scaled + becparknumpercapl2_scaled +
    cnssei_l1ad_scaled,
  ziformula = ~1,
  data = dta,
  family = nbinom2
)
summary(M2_trn_zinb2)
check_collinearity(M2_trn_zinb2)
```

```{r}
#| label: IRR table transportation (full model)
#| code-summary: IRR table transportation (full model)
pred_trn <- ggpredict(
  M2_trn_zinb2,
  terms = "l1adtemp_pw_1_7_mean [all]",
  bias_correction = TRUE
)
pred_trn_df <- as.data.frame(pred_trn) %>%
  rename(
    temperature = x,
    predicted = predicted,
    lower = conf.low,
    upper = conf.high
  )
ref_temp <- pred_knots_trn[[1]]
ref_index <- which.min(
  abs(pred_trn_df$temperature - ref_temp)
)
ref_rate <- pred_trn_df$predicted[ref_index]
pred_trn_df <- pred_trn_df %>%
  mutate(
    IRR = predicted / ref_rate,
    IRR_low = lower / ref_rate,
    IRR_high = upper / ref_rate
  )
temps_interest <- as.numeric(pred_knots_trn)
irr_trn_table <- lapply(
  temps_interest,
  function(temp) {
    pred_trn_df %>%
      slice(
        which.min(abs(temperature - temp))
      )
  }
) %>%
  bind_rows() %>%
  mutate(
    percentile = c(
      "10th percentile",
      "50th percentile",
      "85th percentile"
    ),
    IRR_CI = paste0(
      round(IRR, 2),
      " (",
      round(IRR_low, 2),
      "; ",
      round(IRR_high, 2),
      ")"
    )
  ) %>%
  select(
    percentile,
    temperature,
    IRR_CI
  )
irr_trn_table
```

```{r}
#| label: Curve transportation (full model)
#| code-summary: Curve transportation (full model)
ggplot(
  pred_trn_df,
  aes(
    x = temperature,
    y = IRR
  )
) +
  geom_ribbon(
    aes(
      ymin = IRR_low,
      ymax = IRR_high
    ),
    alpha = 0.20
  ) +
  geom_line(
    linewidth = 1
  ) +
  geom_hline(
    yintercept = 1,
    linetype = "dashed"
  ) +
  geom_vline(
    xintercept = ref_temp,
    linetype = "dashed"
  ) +
  labs(
    x = "Mean temperature (°C)",
    y = "Incidence Rate Ratio (IRR)"
  ) +
  theme_minimal()
```